Comparison of Confidence Interval in the OLS and Ridge Linear Regression Model: A Comparative study via Simulation

### 2024-09-16 draft_1.qmd
updated simulation section

--------\[ aim needs to be updated based on discussion\]

@@

The OLS estimator $\hat{\beta}$ is an unbiased estimator of the ${\beta}$. Under the assumptions od the classical linear regression model the OLS estimator $\hat{\beta}$ is calculated by minimizing the sum of square of the residuals. To measure the variability of the estimate across samples the standard error is calculated as follows, $$ \text{SE}(\hat{\beta}) = \sqrt{\frac{\sigma^2}{\sum (x_i - \bar{x})^2}},
$$ {#eq-2} where, $\sigma^2$ is the variance of the error term. Hence the confidence interval for $\hat{\beta}$ is given as, $$ \hat{\beta} \pm t_{\alpha/2, n-k} \cdot \text{SE}(\hat{\beta})
$$ {#eq-3} where, $t_{\alpha/2, n-k}$ is the critical value for t- distribution with $n - k$ degrees of freedom for $n$ number of samples and $k$ predictors.

@@

------- discuss whether to add different confidence interval

@@

# Simulation Study

>>>>

@@

The OLS estimator $\hat{\beta}$ is an unbiased estimator of the ${\beta}$. Under the assumptions of the classical linear regression model the OLS estimator $\hat{\beta}$ is calculated by minimizing the sum of square of the residuals. To measure the variability of the estimate across samples the standard error is calculated as follows, $$ \text{SE}(\hat{\beta}) = \sqrt{\frac{\sigma^2}{\sum (x_i - \bar{x})^2}},
$$ {#eq-2} where, $\sigma^2$ is the variance of the error term. Hence the confidence interval for $\hat{\beta}$ is given as, $$ \hat{\beta} \pm t_{\alpha/2, n-k} \cdot \text{SE}(\hat{\beta})
$$ {#eq-3} where, $t_{\alpha/2, n-k}$ is the critical value for t- distribution with $n - k$ degrees of freedom for $n$ number of samples and $k$ predictors.

@@

The standard error for Ridge regression is calculated after adjusting for the bias from the regularization parameter, The varinace-covariance matrix of the Ridge Regression estimator is,

$$
\text{Var}(\hat{\beta}_{\text{ridge}}) = \sigma^2 (Z^T Z + \lambda I)^{-1} Z^T Z (Z^T Z + \lambda I)^{-1}
$$ Confidence interval for Ridge Regression is calculated using the Students-t approximation method where t-Statistics is,

$$
t_j = \frac{\hat{\beta}_{\text{ridge}, j}}{\hat{\text{SE}}(\hat{\beta}_{\text{ridge}, j})},
$$ where, $\hat{\text{SE}}(\hat{\beta}_{\text{ridge}, j})$ is the standard error of the j-th coefficient. Hence, the j-th confidence interval is calculated using the following formula, $$
\hat{\beta}_{\text{ridge}, j} \pm t_{\alpha/2, n-p} \cdot \hat{\text{SE}}(\hat{\beta}_{\text{ridge}, j}),
$$ here, $t_{\alpha/2, n-p}$ is the critical value of the student's t-distribution with $n-p$ degrees of freedom.

@@

# Simulation Design

The process of Monte Carlo simulation is used for the simulation study in order to evaluate the performance of the shrinkage parameters. The steps of the procedures is given as follows,

1) The independent variables are generated by following the method of @mcdonald1975monte as follows,

$$
x_{ij} = \left(1 - \rho^2 \right)^{1/2} w_{ij} + \rho w_{i(p+1)}, \quad j = 1, \dots, p, \quad i = 1, \dots, n
$$
here, $w_{ij}$ is the error term $\rho$ is the correlation between two predictors, p is the number of explanatory variables and, n is the number of samples. The number of sample considered for the study are, $n = 30, 100$ and level of collinearity taken are, $\rho = 0.80, 0.85, 0.90, 0.95, 0.99$.

2) Therefore the observations are derive using the following regression model, 
$$
y_i = \beta_1 x_{i1} + \dots + \beta_p x_{ip} + \epsilon_i, \quad i = 1, \dots, n
$$
3) The error term are generated from pseudo-random independent standard normal numbers with mean zero and unit variance.

*** needs to be edited

4) values of s

5) number of iteration

6) calculating of confidence interval

7) calculation of coverage probability

   software used **** needs to be edited
